---
title: Fitting linear models to any matrix representation
author: 
- name: Aaron Lun
  affiliation: Cancer Research UK Cambridge Institute, Cambridge, United Kingdom
date: "Revised: 14 September 2018"
output:
  BiocStyle::html_document:
    toc_float: true
package: flmam 
vignette: >
  %\VignetteIndexEntry{Fitting linear models to any matrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}    
---

```{r, echo=FALSE, results="hide", message=FALSE}
require(knitr)
opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
```

```{r setup, echo=FALSE, message=FALSE}
library(flmam)
set.seed(100)
```

# Introduction

The `r Biocpkg("flmam")` package (pronounced "flim-flam", for the sake of convenience) does exactly what it says on the tin - fit linear models to any matrix.
By taking advantage of `r Biocpkg("beachmat")`, we support any matrix representation including sparse or HDF5-backed matrices.

# Different fitting modes

We support fitting with a design matrix, as in `lm.fit`:

```{r}
y <- matrix(rnorm(20000), ncol=100)
covariate <- runif(ncol(y))
design <- model.matrix(~covariate)
fitLM(y, design)
```

For one-way layouts, greater efficiency can be achieved by passing the factor directly:

```{r}
g <- gl(20, 5)
(fit <- fitLM(y, g))
```

Each row of the output `DataFrame` contains the fit results for the corresponding row of `y`.
By default, a matrix of coefficients is returned in `coefficients`, along with the residual variances in `variances`.

# Different matrix types

For demonstration, let's try it with sparse matrices, namely `dgCmatrix` objects from the `r CRANpkg("Matrix")` package.

```{r}
library(Matrix)
y <- rsparsematrix(2e4, 100, density=0.1)
groups <- gl(20, 5)
fitLM(y, groups)
```

We can also use with HDF5-backed matrices from the `r Biocpkg("HDF5Array")` package.

```{r}
library(HDF5Array)
z <- as(matrix(rnorm(2000000), ncol=100), "HDF5Array")
fitLM(z, groups)
```

# Further tweaking

We can control the rows to be fitted with the `rows` argument.
This is more efficient than subsetting the matrix beforehand.

```{r}
y <- matrix(rnorm(20000), ncol=100)
groups <- gl(10, 10)
fitLM(y, groups, rows=1:10)
```

Users can set `BPPARAM` to specify how parallelization should be achieved.
By default, the registered backend from `bpparam()` will be used.

```{r}
y <- matrix(rnorm(20000), ncol=100)
groups <- gl(10, 10)
fitLM(y, groups, BPPARAM=SnowParam(3))
```

# Session information

```{r}
sessionInfo()
```
